---
title: "569-HW5"
author: "Lian Zuo"
date: "2023-03-04"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment = "")
```

# Q1
## a) the MLE of $\hat\theta$ is 8.1.
## b) the MLE of $\theta^2+2\theta-1$ is 80.81.
## c) the MLE of mean and variance of the random variable Y is 8.1 and 8.1.
## d) the MLE estimate of variance and standard error of $\theta$ is 0.405 and 0.6363961
## e) the  90% Wald’s confidence interval for $\theta$ is [7.053222 , 9.146778].

```{r}
# a
Y <- c(5, 7, 8, 13, 11, 11, 6, 8, 10, 15, 4, 6, 5, 5, 8, 11, 7, 3, 6, 13) 
theta.hat <- sum(Y)/length(Y)
theta.hat
# b
f.theta.hat <- theta.hat^2+2*theta.hat-1
f.theta.hat
# c
mean.y=theta.hat
var.y=theta.hat
list(mean.y,var.y)
# d
var.theta.hat <- theta.hat/length(Y)
se.theta.hat <- sqrt(var.theta.hat)
list(var.theta.hat,se.theta.hat)
#e
CI <- c(theta.hat-qnorm(0.95,0,1)*se.theta.hat,theta.hat+qnorm(0.95,0,1)*se.theta.hat)
CI
```

# Q2
## a)Find an MLE of success probability p is 0.805.
## b) Find the estimated variance of MLE of p is 0.003924375.
## c) Find the estimated standard error of the MLE of p is 0.06264483
## d) Find a 95% Wald confidence interval for the success probability p is (0.6822184 ,0.9277816).
## e) Find an MLE of odds of success p is 4.128205. This odds means that there are about 4 successes for every failure.
```{r}
# a
set.seed(2023)
dat=rbinom(15,40,0.8)
p.hat <- sum(dat)/(15*40)
p.hat
# b
v.p.hat <- p.hat*(1-p.hat)/40
v.p.hat
# c
se.p.hat <- sqrt(v.p.hat)
se.p.hat
# d
CI <- c(p.hat-qnorm(0.975,0,1)*se.p.hat,p.hat+qnorm(0.975,0,1)*se.p.hat)
CI
# e
odds <- p.hat/(1-p.hat)
odds
```

# Q3
## a) The probability that there will be 10 accidents over a 2-year period is 0.1048373
## b) The probability that there will be 3 accidents over an 8-month period is 0.1953668
## c) The probability that there will be at most 4 accidents over a 7-month period is 0.725445
## d) The probability that there will be at least 5 accidents over a 2.5-year period is 0.9991434
## e) The probability that there will be at most 3 accidents over a 1-year period is 0.1512039
```{r}
# a
dpois(10,6*2)
# b
dpois(3,6*(8/12))
# c
ppois(4,6*(7/12))
# d
1-ppois(4,6*2.5)
# e
ppois(3,6)
```

# Q4
## a) Find an MLE $\hat\lambda$ of $\lambda$ is 0.4652241
	
## b) Find the estimated variance of MLE of $\lambda$ is 0.0007190481

## c) Find the estimated standard error of the MLE of $\lambda$ is 0.02681507

## d) Find a 99% Wald confidence interval for the parameter $\lambda$ is (0.3961531 ,0.5342952)
	
## e) Find an MLE of $(\lambda ^2-1)/(2\lambda +1)$ is -0.4058988.

```{r}
#a
x <- c(0,1,2,3,4,5)
y <- c(447,132,42,21,3,2)
mean=sum(x*y)/sum(y)
lamda.hat <- mean
lamda.hat
#b
var.lamda <- lamda.hat/sum(y)
var.lamda
#c
se.lamda <-sqrt(var.lamda)
se.lamda
#d
CI <- c(lamda.hat-qnorm(0.99+0.01/2)*se.lamda,lamda.hat+qnorm(0.99+0.01/2)*se.lamda)
CI
#e
(lamda.hat^2-1)/(2*lamda.hat+1)
```


